Follow the Flow of Rain and Paint: \break An Output Sensitive Contour Tree Algorithm

### 2014-09-22 paint-contour-tree.tex
working on upper bounds

\begin{definition}
\label{def:path} A \emph{path decomposition}, $P(T)$, is a partition of $T$ into 
 a set of vertex disjoint leaf paths.  Note that the number of paths in $P(T)$ must equal the number of leaves in $T$.

 The \emph{cost} of $P(T)$ is $\cost(P(T)) = \sum_{p\in P(T)} |p| \log |p|$. (We use $|p|$ to denote the number of vertices in $p$.)
\end{definition}

@@

Our main theorem is the following.

\begin{theorem} \label{thm:runtime} $\sum_{v \in T} \log |H_v| = O(\cost(\pmax(T)))$.
\end{theorem}

@@

\begin{lemma}
\label{lem:pathBounds} If $p$ is heavy, $\sum_{v\in p} \log |H_v| = O(|p| \log |p|)$.
If $p$ is light, $\sum_{v\in p} \log |H_v| = O(|p| \log |H_p|)$.
% Let T be a binary tree and let $\chi(T)$ and $P(T)$ be a valid coloring and path decomposition of $T$, respectively.
% Then, 
% \begin{compactenum}
%  \item for any heavy path $p\in P(T)$, $\sum_{v\in p} \log |H_v| = O(|p| \log |p|)$.
%  \item for any light path $p\in P(T)$, $\sum_{v\in p} \log |H_v| = O(|p| \log |H_{p}|)$.
% \end{compactenum}
\end{lemma}
\begin{proof} \ignore{
 By $\Obs{decrease}$ we know that for any vertex $v\in p$, $|H_v|\leq |H_{p}| + |p|$ (as $v$ is a descendant of $r_p$ along $p$).
}
For $v \in p$, $|H_v|\leq |H_{p}| + |p|$ (as $v$ is a descendant of $b(p)$ along $p$).
Hence, $\sum_{v\in p} \log |H_v| \leq \sum_{v\in p} \log(|H_{p}| + |p|)  = |p| \log (|H_{p}| + |p|)$.
 If $p$ is a heavy path, then $|p| \log (|H_{p}| + |p|) = O(|p| \log |p|)$. If $p$ is light, then 
 $|p| \log (|H_{p}| + |p|) = O(|p| \log |H_{p}| )$.
\end{proof}

@@

\begin{definition} \label{def:support} A light path $p$ is called \emph{supported} if: 
at least $|H_p|/100$ vertices in $p$ lie in heavy ancestors $q$ such that $|H_q| > |H_p|/100$.
%
%
% there exists
%a contiguous (from $p$) chain of ancestors $q_1, q_2, \ldots, q_k$ (in $P(T)$) such that $|H_{q_k}| \leq 99|H_p|/100$
%and all $q_i$'s are heavy.
\end{definition}

Let $\cL$ be the set of light paths, $\cL'$ be the set of supported light paths, and $\cH$ be the set of heavy paths.
%We have a claim that also bounds heap sizes for special set of paths.

\begin{claim} \label{clm:above} $\sum_{p \in \cL'} |H_p| \log |H_p| = O(\sum_{p \in \cH} |p|\log |p|)$.
\end{claim}

\begin{proof} Pick $p \in \cL'$. As we traverse up $q_1, q_2, \ldots, q_k$, there
are certain vertices of $H_p$ that lie in these paths. There must be at least $|H_p|/100$
such vertices, since $H_{q_k} \leq 99H_p/100$. These are encountered in
a fixed order. Let $H'_p$ be the first half of these vertices. For all $v \in H'_p$
$|H_v| \geq |H_p|/200$. Hence, $|H_p|\log |H_p| = O(\sum_{v \in H'_p} \log |H_v|)$
and $\sum_{p \in L'} |H_p|\log |H_p| = O(\sum_{p \in L'} \sum_{v \in H'_p} \log |H_v|)$.

Pick another $\widetilde{p} \in \cL'$. If $p$ and $\widetilde{p}$ are ancestor-descendant,
then their corresponding supporting ancestor chains are disjoint. Any $v \in H'_p$
can only occur in one other $H'_q$ (for $q \in \cL'$), because $v$ has exactly two colors.
In other words, any $v$ appears at most twice in the above double summation.
These vertices are all present in heavy paths, so we can bound
the double summation by $\sum_{p \in \cH} \sum_{v \in p} \log |H_v|$.
%
%Pick $q \in Q$ and some $v \in H_q$. As we traverse up the tree passing
%through $A(q)$, we must encounter some $p \in A(q)$ such that $v \in p$. Furthermore,
%there is an order in which these vertices (of $H_q$) occur. Let $H'$ be the first half
%of vertices that occur. For all $v \in H'$, $|H_v| \geq |H_q|/2$. Hence,
%$|H_q| \log |H_q| \leq 4 \sum_{v \in H'} \log |H_v| \leq 4 \sum_{v \in H_q} |H_v|$.
%So $\sum_{q \in Q} |H_q| \log |H_q| \leq 4 \sum_{q \in Q} \sum_{v \in H_q} |H_v|$.
%
%Since nodes in $Q$ are incomparable, all the $H_q$'s (for $q \in Q$) are disjoint.
%Each such $v$ must occur in some path $p \in A(Q)$. Thus,
%$\sum_{q \in Q} |H_q| \log |H_q| \leq 4 \sum_{p \in A(Q)} \sum_{v \in p} |H_v|$.
We apply the bound of \Lem{pathBounds} to complete the proof.
\end{proof}

Now, for the critical piece of the approach. Focus on $\pmax(T)$. Let $L$ be the set of light paths, 
and $\hat{L} \subseteq L$ be the set of paths such that no ancestor of $p \in L'$
is light. Obviously, $\hat{L}$ is an anti-chain in $P(T)$.

@@

The main challenge is proving \Lem{light}, which we defer to the next section.

\ignore{
\begin{definition}
 Let $\chi(T)$ be a valid coloring of a binary tree $T$.  We define a set of disjoint paths over the vertices of 
 $T$ as follows.  
 The corresponding collection of paths that all such edges correspond to is a path decomposition of $T$, which 
 we refer to as a \emph{maximum} path decomposition, and is denoted by $\mathcal{P}(T)$.  
\end{definition}
}

Our focus is on $\pmax(T)$.
We can simplify a little. We split the summation and apply \Lem{pathBounds}. Finally,
we note that for $p \in \cL$, $|p| \leq |H_p|$.
\begin{eqnarray*} 
\sum_{p \in \cL} \sum_{v \in p} \log |H_v| & = & \sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| + \sum_{p \in \cL'} \sum_{v \in p} \log |H_v|\\
& = & \sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| + O(\sum_{p \in \cL'} |p|\log |H_p|) = \sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| + O(\sum_{p \in \cL'} |H_p|\log |H_p|)
\end{eqnarray*}

\ignore{
From the previous section we now have a lower bound of $\Omega(f_{path}(P))$ for computing the merge tree in terms of any path decomposition $P$. 
We now show that the running time of our algorithm for computing the merge tree is upper bounded by this expression for a specific path decomposition, 
which we will call the maximum path decomposition (defined below).
Specifically, for a given valid coloring $\chi$ of a tree $T$, we will prove that for the maximum path decomposition $f_{alg}(\chi(T)) = O(f_{path}(P(T)))$. 
%using the observations about path decompositions from 
%\Sec{background}, which will then imply $f_{alg}(\chi(T)) = \Theta(f_{path}(P(T)))$.

@@

We now construct the shrub of unsupported light paths. Consider $p \in \cL \setminus \cL'$,
and traverse the chain of ancestors from $p$. Eventually, we must reach another light path $q$.
(If not, we have reached the root $r$ of $\pmax(T)$. But $|H_r| = 0$, implying $p$ is supported.)
Insert edge from $p$ to $q$, so $q$ is the parent of $p$ in $\cU$. This construction leads
to the shrub forest of $\cL \setminus \cL'$, where every unsupported light path is an internal node, and all
roots are supported light paths. \Lem{support} is a direct consequence of the following,
which is the really the main lemma.

\begin{proof} (of \Lem{support}) We have $\sum_{p \in \cU} \sum_{v \in p} \log |H_v| = O((\log |H_r|) \sum_{p \in \cU} |p|) = O(|H_r|\log |H_r|)$.
We sum this over all $\cU$ in the shrub forest, and note that roots in the shrub forest are supported light paths.
\end{proof}

@@

\Lem{cu-root} asserts the root $r$ in $\cU$ pretty much encompasses all sizes and heaps in $\cU$.
Consider $q \in \cL \setminus \cL'$ that is a child of $p$ in the shrub $\cU$.
By construction, there is a heavy path (denoted $h(q)$) that is a child of $p$ in $\pmax(T)$,
and an ancestor of $q$. Furthermore, $|H_{h(q)}| > 99|H_q|/100$.
We begin with a key property, which is where the construction of $\pmax(T)$ enters the picture.

@@

\begin{lemma}
\label{lem:geometric}
 Let $p$ be the parent of $q$ in the shrub $\cU$. Then $|H_p|\geq \frac{3}{2}|H_q|$. 
\end{lemma}
\begin{proof} We know that $h(q)$ is a child of $p$ in $\pmax(T)$. Consider the base
of $h(q)$, which is a node $w$ in $T$. Let $v$ denote the sibling of $w$ in $T$.
Their parent is called $u$. Note that both $u$ and $v$ are nodes in the path $p$.
Now, the decomposition $\pmax(T)$ put $u$ and $v$ in the same path $p$. This
implies $|H_v| \geq |H_w|$. Since $|H_u| \geq |H_v| + |H_w| - 1$,
$|H_u| \geq 2|H_w| - 1$. Let $b$ be the base of $p$.
We have $|H_p| = |H_b| \geq |H_u| - |p| \geq 2|H_w| - |p| - 2$. Since $p$ is a light path, $|p| < \sqrt{|H_p|}/100$.
Applying this bound, we get $|H_p| \geq (2-\delta)|H_w|$ (for small constant $\delta > 0$).
Since $w$ is the base of $h(q)$, $H_w = H_{h(q)}$. We apply the bound $|H_{h(q)}| > 99|H_q|/100$,
to complete the proof.

@@

This immediately proves part (i) of \Lem{cu-root}. Part (ii) requires some more work.

@@

>>>>

\begin{definition}
\label{def:path} A \emph{path decomposition}, $P(T)$, is a partition of $T$ into 
 a set of vertex disjoint leaf paths.  Note that the number of paths in $P(T)$ must equal the number of leaves in $T$.
% 
% The \emph{cost} of $P(T)$ is $\cost(P(T)) = \sum_{p\in P(T)} |p| \log |p|$. (We use $|p|$ to denote the number of vertices in $p$.)
\end{definition}

@@

Our main theorem is the following. We use $|p|$ to denote the number of vertices in $p$.

\begin{theorem} \label{thm:runtime} $\sum_{v \in T} \log |H_v| = O(\sum_{p\in P(T)} |p| \log |p|)$.
\end{theorem}

@@

\begin{lemma}
\label{lem:pathBounds} If $p$ is heavy, $\sum_{v\in p} \log |H_v| = O(|p| \log |p|)$.
If $p$ is light, $\sum_{v\in p} \log |H_v| = O(|H_p| \log |H_p|)$.
% Let T be a binary tree and let $\chi(T)$ and $P(T)$ be a valid coloring and path decomposition of $T$, respectively.
% Then, 
% \begin{compactenum}
%  \item for any heavy path $p\in P(T)$, $\sum_{v\in p} \log |H_v| = O(|p| \log |p|)$.
%  \item for any light path $p\in P(T)$, $\sum_{v\in p} \log |H_v| = O(|p| \log |H_{p}|)$.
% \end{compactenum}
\end{lemma}
\begin{proof} \ignore{
 By $\Obs{decrease}$ we know that for any vertex $v\in p$, $|H_v|\leq |H_{p}| + |p|$ (as $v$ is a descendant of $r_p$ along $p$).
}
For $v \in p$, $|H_v|\leq |H_{p}| + |p|$ (as $v$ is a descendant of $b(p)$ along $p$).
Hence, $\sum_{v\in p} \log |H_v| \leq \sum_{v\in p} \log(|H_{p}| + |p|)  = |p| \log (|H_{p}| + |p|)$.
 If $p$ is a heavy path, then $|p| \log (|H_{p}| + |p|) = O(|p| \log |p|)$. If $p$ is light, then 
 $|p| \log (|H_{p}| + |p|) = O(|p| \log |H_{p}| )$. For light paths, $|p| = O(|H_p|)$.
\end{proof}

@@

\begin{definition} \label{def:support} A light path $p$ is \emph{supported} if
at least $|H_p|/100$ vertices $v$ in $H_p$ lie in heavy ancestors.
%
%
% there exists
%a contiguous (from $p$) chain of ancestors $q_1, q_2, \ldots, q_k$ (in $P(T)$) such that $|H_{q_k}| \leq 99|H_p|/100$
%and all $q_i$'s are heavy.
\end{definition}

Let $\cL$ be the set of light paths, $\cL'$ be the set of supported light paths, and $\cH$ be the set of heavy paths
given by $\pmax(T)$.
We now construct the shrub of unsupported light paths. Consider $p \in \cL \setminus \cL'$,
and traverse the chain of ancestors from $p$. Eventually, we must reach another light path $q$.
(If not, we have reached the root $r$ of $\pmax(T)$. Hence, $p$ is supported.)
Insert edge from $p$ to $q$, so $q$ is the parent of $p$ in $\cU$. This construction leads
to the shrub forest of $\cL \setminus \cL'$, where every unsupported light path is an internal node, and all
roots are supported light paths.

Most of work goes into proving the following technical lemma.

We split the remaining argument into two subsections. We first prove \Thm{runtime} from \Lem{cu-root},
which mostly imvolves routing calculations. Then we prove \Lem{cu-root}, where most
of the interesting work happens.

\subsection{Proving \Thm{runtime}} \label{sec:thm-runtime}

We split the summation into heavy, light, and unsupported light paths.
\begin{eqnarray*} 
\sum_{p \in \cL} \sum_{v \in p} \log |H_v| & = & \sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| + \sum_{p \in \cL'} \sum_{v \in p} \log |H_v| + \sum_{p \in \cH} \sum_{v \in p} \log |H_v|
%& = & \sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| + O(\sum_{p \in \cL'} |p|\log |H_p|) = \sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| + O(\sum_{p \in \cL'} |H_p|\log |H_p|)
\end{eqnarray*}
The last term can be bounded by $O(\sum_{p \in \pmax(T)} |p|\log |p|)$, by \Lem{pathBounds}.
The second term can be bounded by $O(\sum_{p \in \cL'} |H_p| \log |H_p|)$, by \Lem{pathBounds} again.
The following claim shows that in turn is at most the last term.

%We have a claim that also bounds heap sizes for special set of paths.

\begin{claim} \label{clm:above} $\sum_{p \in \cL'} |H_p| \log |H_p| = O(\sum_{q \in \cH} \sum_{v \in q} \log |H_v|)$.
\end{claim}

\begin{proof} 
Pick $p \in \cL'$. 
%Let the first $|H_p|/200$ vertices the lie in heavy paths be denoted as $H'_p$. 
%Observe that for any $v \in H'_p$, 
%We have $|H_p| \log |H_p| = O(\sum_{v \in H'_p} \log |H_p|)$
%$= O(\sum_{v \in H'_p} \log |H_v|)$. Since all the vertices lie in heavy paths,
%we can write this as $O(\sum_{q \in \cH} \sum_{v \in H'_p \cap q} \log |H_v|)$.
%Summing over all $p$, the expression is $\sum_{q \in \cH} \sum_{p \in \cL'} \sum_{v \in H'_p \cap q} \log |H_v|$.
%
%
As we traverse the heavy ancestors, there
are at least $|H_p|/100$ vertices of $H_p$ that lie in these paths. These are encountered in
a fixed order. Let $H'_p$ be the first $|H_p|/200$ of these vertices. When $v \in H'_p$
is encountered, there are $|H_p|/200$ vertices of $H_p$ not yet encountered. Hence,
$|H_v| \geq |H_p|/200$. Hence, $|H_p|\log |H_p| = O(\sum_{v \in H'_p} \log |H_v|)$.
Since all the vertices lie in heavy paths,
we can write this as $O(\sum_{q \in \cH} \sum_{v \in H'_p \cap q} \log |H_v|)$.
Summing over all $p$, the expression is $\sum_{q \in \cH} \sum_{p \in \cL'} \sum_{v \in H'_p \cap q} \log |H_v|$.
%and $\sum_{p \in \cL'} |H_p|\log |H_p| = O(\sum_{p \in \cL'} \sum_{v \in H'_p} \log |H_v|)$.

Pick another $\widetilde{p} \in \cL'$. If $p$ and $\widetilde{p}$ are ancestor-descendant,
then their corresponding supporting ancestor chains are disjoint. Any $v \in H'_p$
can only occur in one other $H'_{\hat{p}}$ (for $\hat{p} \in \cL'$), because $v$ has exactly two colors.
In other words, any $\log |H_v|$ appears at most twice in the above triple summation.
Hence, we can bound it by $O(\sum_{q \in \cH} \sum_{v \in q} \log |H_v|)$.
%
%Pick $q \in Q$ and some $v \in H_q$. As we traverse up the tree passing
%through $A(q)$, we must encounter some $p \in A(q)$ such that $v \in p$. Furthermore,
%there is an order in which these vertices (of $H_q$) occur. Let $H'$ be the first half
%of vertices that occur. For all $v \in H'$, $|H_v| \geq |H_q|/2$. Hence,
%$|H_q| \log |H_q| \leq 4 \sum_{v \in H'} \log |H_v| \leq 4 \sum_{v \in H_q} |H_v|$.
%So $\sum_{q \in Q} |H_q| \log |H_q| \leq 4 \sum_{q \in Q} \sum_{v \in H_q} |H_v|$.
%
%Since nodes in $Q$ are incomparable, all the $H_q$'s (for $q \in Q$) are disjoint.
%Each such $v$ must occur in some path $p \in A(Q)$. Thus,
%$\sum_{q \in Q} |H_q| \log |H_q| \leq 4 \sum_{p \in A(Q)} \sum_{v \in p} |H_v|$.
%We apply the bound of \Lem{pathBounds} to complete the proof.
\end{proof}

The first term (unsupported light paths) can be charged to the second term (supported light paths).
This is where the critical \Lem{cu-root} plays a role.

\begin{claim} \label{clm:first} $\sum_{p \in \cL \setminus \cL'} \sum_{v \in p} \log |H_v| = O(\sum_{p \in \cL'} |H_p| \log |H_p|)$.
\end{claim}

\begin{proof} Let $\cU$ denote a connected component of the shrub forest.
We have $\sum_{p \in \cL \setminus \cL'} \leq \sum_{\cU} \sum_{p \in \cU} \sum_{v \in p} \log |H_v|$.
By \Lem{cu-root}, $|H_v| = O(|H_r|)$, where $r$ is the root of $\cU$.
Furthermore, $\sum_{p \in \cU} |p| = O(|H_r|)$. 
We have $\sum_{p \in \cU} \sum_{v \in p} \log |H_v| = O((\log |H_r|) \sum_{p \in \cU} |p|) = O(|H_r|\log |H_r|)$.
We sum this over all $\cU$ in the shrub forest, and note that roots in the shrub forest are supported light paths.
\end{proof}

\ignore{
Now, for the critical piece of the approach. Focus on $\pmax(T)$. Let $L$ be the set of light paths, 
and $\hat{L} \subseteq L$ be the set of paths such that no ancestor of $p \in L'$
is light. Obviously, $\hat{L}$ is an anti-chain in $P(T)$.

@@

The main challenge is proving \Lem{light}, which we defer to the next section.
\begin{definition}
 Let $\chi(T)$ be a valid coloring of a binary tree $T$.  We define a set of disjoint paths over the vertices of 
 $T$ as follows.  
 The corresponding collection of paths that all such edges correspond to is a path decomposition of $T$, which 
 we refer to as a \emph{maximum} path decomposition, and is denoted by $\mathcal{P}(T)$.  
\end{definition}

From the previous section we now have a lower bound of $\Omega(f_{path}(P))$ for computing the merge tree in terms of any path decomposition $P$. 
We now show that the running time of our algorithm for computing the merge tree is upper bounded by this expression for a specific path decomposition, 
which we will call the maximum path decomposition (defined below).
Specifically, for a given valid coloring $\chi$ of a tree $T$, we will prove that for the maximum path decomposition $f_{alg}(\chi(T)) = O(f_{path}(P(T)))$. 
%using the observations about path decompositions from 
%\Sec{background}, which will then imply $f_{alg}(\chi(T)) = \Theta(f_{path}(P(T)))$.

@@

@@

\Lem{cu-root} asserts the root $r$ in $\cU$ pretty much encompasses all sizes and heaps in $\cU$.
%Consider $q \in \cL \setminus \cL'$ that is a child of $p$ in the shrub $\cU$.
We begin with a key property, which is where the construction of $\pmax(T)$ enters the picture.

@@

\begin{lemma}
\label{lem:geometric}
 Let $p$ be the parent of $q$ in the shrub $\cU$. Then $|H_p|\geq \frac{3}{2}|H_q|$. 
\end{lemma}
\begin{proof} By construction, there is a heavy path (denoted $h(q)$) that is a child of $p$ in $\pmax(T)$,
and an ancestor of $q$. The chain of ancestors from $q$ to $h(q)$ consists only of heavy paths.
These paths contain at most $|H_p|/100$ vertices of $H_p$, so $|H_{h(q)}| \geq 99|H_p|/100$.

Consider the base
of $h(q)$, which is a node $w$ in $T$. Let $v$ denote the sibling of $w$ in $T$.
Their parent is called $u$. Note that both $u$ and $v$ are nodes in the path $p$.
Now, the decomposition $\pmax(T)$ put $u$ and $v$ in the same path $p$. This
implies $|H_v| \geq |H_w|$. Since $|H_u| \geq |H_v| + |H_w| - 1$,
$|H_u| \geq 2|H_w| - 1$. Let $b$ be the base of $p$.
We have $|H_p| = |H_b| \geq |H_u| - |p| \geq 2|H_w| - |p| - 2$. Since $p$ is a light path, $|p| < \sqrt{|H_p|}/100$.
Applying this bound, we get $|H_p| \geq (2-\delta)|H_w|$ (for small constant $\delta > 0$).
Since $w$ is the base of $h(q)$, $H_w = H_{h(q)}$. We apply the bound $|H_{h(q)}| > 99|H_q|/100$,
to complete the proof.

@@

This immediately proves part (i) of \Lem{cu-root}. Part (ii) requires much more work.
It is helpful to spell out how the various heaps evolve. Consider $p \in \cU$,
which has children $q_1, q_2, \ldots, q_k$ in $\pmax(T)$.
Furthermore, as a leaf path in $p$, some leaf
$\ell$ of $T$ in one of the ends of $p$. The heap $H_p$ is the
union of $H_\ell$ and $H_{q_1}, H_{q_2}, \ldots, H_{q_k}$, with the vertices in $p$
removed. Note that some of the $q_i$'s are the corresponding $h(q')$ vertices,
for a child $q'$ in $\cU$.

@@

We now construct the shrub of unsupported light paths. Consider $p \in \cL \setminus \cL'$,
and traverse the chain of ancestors from $p$. Eventually, we must reach another light path $q$.
(If not, we have reached the root $r$ of $\pmax(T)$. Hence, $p$ is supported.)
Insert edge from $p$ to $q$, so $q$ is the parent of $p$ in $\cU$. This construction leads
to the shrub forest of $\cL \setminus \cL'$, where every unsupported light path is an internal node, and all
roots are supported light paths.

We list out some important claims about this shrub forest.
We focus on $\cU$, a connected component (shrub) in the shrub forest and denote the 
the root by $r(\cU)$.

\begin{claim} \label{clm:loss} Consider $p \in \cL \setminus \cL'$ that is in $\cU$. Let $Q$
be the set of heavy ancestors of $p$ that are descendants of $r(\cU)$. Then $\sum_{q \in Q} |q \cap H_p| \leq |H_p|/50$.
\end{claim}

\begin{proof} Suppose not, for contradictions sake. As we traverse the heavy ancestors of $p$,
we encounter various vertices of $H_p$ in the ancestors $q$. Consider the first $|H_p|/100$
such vertices of $p$ encountered. At each such $v$, there are at least $|H_p|/100$
vertices \emph{not yet} encountered, so $|H_v| > |H_p|/100$. Each of these vertices $v$
is supported, implying that $p$ is a supported light path.
\end{proof}

\Lem{support} is a direct consequence of the following,
which is the really the main lemma.

\begin{proof} (of \Lem{support})
\end{proof}
